Information Disclosure Concerns\\in The Age of Wearable Computing

### 2016-02-02 USEC2016-camera-ready/camera-ready.tex
Made title wrap better
used official IEEE template (rather than implicit)

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Information Disclosure Concerns in The Age of Wearable Computing}

@@

\subsection{Survey Questions}
\noindent We report on participants' responses to 23 questions across 4 survey sections:  \\

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item 2 reading comprehension questions
\item 6 questions regarding possible wearables scenarios
\item 1 open-ended wearables risk question
\item 14 demographic questions \\
\end{itemize}

To reduce fatigue, we gave our participants a randomly selected subset of wearables scenarios. The average survey completion time was 11.5 minutes, which included four questions that we omitted from this paper due to lack of participants' familiarity with specific devices and a misguided attempt to directly compare smartphones and wearables. See the appendix for details.\\

\noindent {\bfseries Comprehension Questions}
Because participants might be biased to specific companies (e.g., visceral reactions to Google Glass based on popular media stories), we framed our scenarios around a fictitious wearable. The survey introduced participants to the ``Cubetastic3000,'' which was the basis for all questions on wearables risks. We highlighted the capabilities of this device and described use cases:\\

@@

To guarantee that participants understood its capabilities, we asked two multiple-choice comprehension questions and removed participants who did not answer both correctly.\\

\noindent {\bfseries Wearable Scenarios}
We presented scenarios involving data captured by the Cubetastic3000 and asked participants to rate how upset they would be if a particular type of data (e.g., how much you exercise) was shared without permission with a particular recipient (e.g., work contacts). The purpose for using this question format was to determine how upset participants would be if data were inappropriately disclosed, and the extent to which their reactions were based on the data type and recipient. Responses were reported on a 5-point Likert scale (from ``indifferent'' to ``very upset''). Figure \ref{fig:prompt} shows an example.  Specifically, questions were of the form: \\

\begin{quotation}
\noindent
\textit{``How would you feel if an app on your Cubetastic3000 learned (data) and shared it with (recipient), without asking you first?''}\\
\end{quotation}

@@

We combined 72 data types ({\it data}) with 4 recipients ({\it recipient}) to form 288 scenarios (Table \ref{all-vur}). Each participant answered 6 randomly drawn questions, displayed in random order. We clarified that ``app'' meant that the data was not shared with anyone else but a server. \\

@@

\section{Results}
We had at least 141 responses per data type, 2,779 per recipient, and 35 responses per each unique data type/recipient combination. In this section, we present participants' responses to the various data disclosure scenarios and discuss how various factors contributed to their risk perceptions. We conclude with self-reported general wearables concerns.\\

@@

Although certain data is considered unanimously upsetting to have shared, it is interesting to note that no data was considered unanimously non-upsetting to have shared, nor were there any data types that evoked strong disagreement between participants (i.e., bimodal). Generally, the average concern magnitude was inversely correlated with the standard deviation, which suggests the presence of ceiling effects for the most concerning data types. For the complete ranked list of data types in this study, see Table \ref{all-vur}.\\

@@

We note that this chi-square test violates the assumption of independent observations, since participants responded to multiple scenarios with multiple recipients. But based on the randomization of treatments and large sample size, we do not believe that this significantly impacted our results. Similarly, we are unaware of a more appropriate test (beyond using mixed effects modeling), given our data format. Cochran's Q requires binary outcomes (i.e., participants would have had to answer only one question for each data recipient, preventing us from adequately controlling for data type) and a repeated measures ANOVA requires normality (our data was not normally distributed). Nonetheless, we repeated our analysis using one randomly-selected data point per participant and found that our selected test was robust to this violation. Therefore, we conclude that participants were significantly more concerned about having their data seen by a human than an application, though differences between human groups such as the public, friends, and work contacts were not as significant. Our results motivate mechanisms for data taint tracking and accidental data sharing prevention.

However, we do not claim that there are no distinctions between the friends, public, and work contact recipients. People are more comfortable sharing certain data types with certain human recipients. For instance, participants were significantly more uncomfortable sharing if they were lying, nervous, or stressed to work contacts compared to the rest of the data recipients. We believe a more fine-grained study is required to measure these effects accurately, although we have insight into the existence of these trends. Table \ref{all-vur} shows the complete VURs and rankings of all data types by recipient.\\

@@

\noindent
\textit{P1256: ``They can be hacked and then your security will be compromised.''}\\
\end{quotation}

Other common concerns included being unaware of what the device is collecting, doing, or which information it is using (Being Unaware), long-term health effects caused from wearing the device such as cancer from EM waves (Health), and safety hazards from wearing the device, such as distractions that cause car accidents (Safety).\\

@@

\noindent
\textit{P1038: ``Becoming distracted by the devices while doing other activities that require concentration such as driving.''}\\

@@

Interestingly, a few participants were concerned with resulting changes in social behaviors, such as dependence on devices or spending less time with loved ones (Social Impact). \\

\begin{quotation}
\noindent
\textit{P1425: I think the biggest risk is how they may effect society as a whole... a wearable technology that's always on and available may push things even further to the point where people spend less time actually interacting with loved ones, and applying their own critical thinking in certain situations, instead always relying on their devices.}\\
\end{quotation}

The landscape of users' perceived risks associated with wearables is broad, encompassing concepts outside of privacy and security. We therefore hope this motivates other researchers to investigate these other risks. \\

@@

While we initially observed an effect on VURs based on whether or not participants claimed to already own wearables (57.0\% vs. 60.8\%, respectively; Mann-Whitney $U=202,896$, $p\textless0.032$), this difference did not remain significant upon correcting for multiple testing (Bonferroni corrected $\alpha=0.01$). The effect of a participant's gender also did not remain significant upon correcting for multiple testing. We observed no correlation between a participant's education level and VUR.\\

@@

While the first three categories are self-explanatory, the latter three categories are based on user characteristics. We defined {\it behavioral information} as observations about the user's activities; {\it biometric information} as measurements of the user's body; and {\it demographic information} as non-biometric information about the user's traits.

The risks for data types fell into five categories:

@@

Although we presented our participants with a prompt illustrating all the benefits of a wearable device, the questions we asked isolated the risk from the benefits of sharing data with a particular recipient. We sacrificed context due to the complexity of the question necessary for the participant to answer correctly. Users are willing to tolerate risks if there is enough benefit associated with that risk. We do believe that since all of our questions were out of context, our study does represent what data people unambiguously would like to be private and secure, and is accurate for measuring user perceptions of wearables.\\

@@

\section{Related Work}
\noindent {\bfseries User Perceptions} 
While risk communication for the physical world has been examined for several decades (e.g.,~\cite{slovic2000perception, Fischhoff,Morgan2001}), research into effectively communicating computer-based risks has only recently been researched. For example, both Garg {\it et al.}\ and Blythe {\it et al.}\ show that due to varying perceptions and abilities that correlate with demographic factors, computer-based risk communication should employ some degree of demographic targeting~\cite{Garg2012,Blythe2011}. While this work is likely applicable to wearable computing risk communication, we believe that a better understanding of users'  information risk acceptance in this domain is warranted prior to examining risk communication.

@@

One limitation of user perceptions is that people do not always have enough information to make privacy-sensitive decisions. Even if users did have this information, it has been shown that users often trade off long-term privacy for short-term benefits~\cite{acquisti2005privacy}. Furthermore, actual behavior may deviate from stated privacy preferences~\cite{spiekermann2001privacy}. However, understanding user concerns is a necessary first step not only for risk communication, but preventative measures against breaches of privacy and security in a new threat landscape. \\

\noindent {\bfseries Ubiquitous Sensing} We are rapidly moving towards a world of ubiquitous sensing and data capture, with ensuing privacy challenges~\cite{abowd2000charting,palen2003unpacking,camp2000internet}. Roesner {\it et al.}\ urge the community to address potential concerns for wearable devices before the technologies become widespread~\cite{roesner2014security} and explore the unique problems present in terms of law and policy~\cite{roesner2014augmented}. Privacy preservation research in this age of ubiquitous sensing include frameworks to design for privacy~\cite{bellotti1993design,camp2003designing,langheinrich2001privacy}, protocols for anonymous communication~\cite{cornelius2008anonysense}, evaluation metrics for privacy~\cite{scholtz2004toward}, and privacy models~\cite{hong2004privacy, jiang2002approximate}. Our work aims to guide these efforts with an insight into user acceptance of common information risks. \\

\noindent {\bfseries Lessons from Smartphones} Not long ago, smartphones revolutionized applications' access to  data. While this tends to benefit users, they often do not think of the privacy implications. There are still many unresolved concerns such as the opaqueness that prevents users from fully understanding how applications are using their data or rogue applications inappropriately accessing data~\cite{1_kane_2010, zhou2011taming}, which are applicable to wearables.

Felt {\it et al.} previously studied the security concerns of smartphone users by conducting a large-scale online survey~\cite{Felt}. Their survey asked 3,115 smartphone users about 99 risk scenarios. Participants were asked how upset they would be if a certain action occurred without their permission. Participants rated each situation on a Likert scale ranging from ``indifferent (1)'' to ``very upset (5).'' Our methodology closely follows that study, but with scenarios chosen to shed light on the security and privacy risks of wearable devices.\\

@@

\input{tex-inputs/appendixes.tex}
\input{tex-inputs/VURcombined.tex}

>>>>

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Information Disclosure Concerns\\in The Age of Wearable Computing}

@@

\subsection{Survey Questions}
\noindent We report on participants' responses to 23 questions across 4 survey sections:

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item 2 reading comprehension questions
\item 6 questions regarding possible wearables scenarios
\item 1 open-ended wearables risk question
\item 14 demographic questions 
\end{itemize}

To reduce fatigue, we gave our participants a randomly selected subset of wearables scenarios. The average survey completion time was 11.5 minutes, which included four questions that we omitted from this paper due to lack of participants' familiarity with specific devices and a misguided attempt to directly compare smartphones and wearables. See the appendix for details.

\noindent {\bfseries Comprehension Questions}
Because participants might be biased to specific companies (e.g., visceral reactions to Google Glass based on popular media stories), we framed our scenarios around a fictitious wearable. The survey introduced participants to the ``Cubetastic3000,'' which was the basis for all questions on wearables risks. We highlighted the capabilities of this device and described use cases:

@@

To guarantee that participants understood its capabilities, we asked two multiple-choice comprehension questions and removed participants who did not answer both correctly.

\noindent {\bfseries Wearable Scenarios}
We presented scenarios involving data captured by the Cubetastic3000 and asked participants to rate how upset they would be if a particular type of data (e.g., how much you exercise) was shared without permission with a particular recipient (e.g., work contacts). The purpose for using this question format was to determine how upset participants would be if data were inappropriately disclosed, and the extent to which their reactions were based on the data type and recipient. Responses were reported on a 5-point Likert scale (from ``indifferent'' to ``very upset''). Figure \ref{fig:prompt} shows an example.  Specifically, questions were of the form:

\begin{quotation}
\noindent
\textit{``How would you feel if an app on your Cubetastic3000 learned (data) and shared it with (recipient), without asking you first?''}
\end{quotation}

@@

We combined 72 data types ({\it data}) with 4 recipients ({\it recipient}) to form 288 scenarios (Table \ref{all-vur}). Each participant answered 6 randomly drawn questions, displayed in random order. We clarified that ``app'' meant that the data was not shared with anyone else but a server.

@@

\section{Results}
We had at least 141 responses per data type, 2,779 per recipient, and 35 responses per each unique data type/recipient combination. In this section, we present participants' responses to the various data disclosure scenarios and discuss how various factors contributed to their risk perceptions. We conclude with self-reported general wearables concerns.

@@

Although certain data is considered unanimously upsetting to have shared, it is interesting to note that no data was considered unanimously non-upsetting to have shared, nor were there any data types that evoked strong disagreement between participants (i.e., bimodal). Generally, the average concern magnitude was inversely correlated with the standard deviation, which suggests the presence of ceiling effects for the most concerning data types. For the complete ranked list of data types in this study, see Table \ref{all-vur}.

@@

We note that this chi-square test violates the assumption of independent observations, since participants responded to multiple scenarios with multiple recipients. But based on the randomization of treatments and large sample size, we do not believe that this significantly impacted our results. Similarly, we are unaware of a more appropriate test (beyond using mixed effects modeling), given our data format. Cochran's Q requires binary outcomes (i.e., participants would have had to answer only one question for each data recipient, preventing us from adequately controlling for data type) and a repeated measures ANOVA requires normality (our data was not normally distributed). Nonetheless, we repeated our analysis using one randomly-selected data point per participant and found that our selected test was robust to this violation. Therefore, we conclude that participants were significantly more concerned about having their data seen by a human than an application; differences between human recipients were not as significant. Our results motivate mechanisms for tracking data flow and preventing accidental data sharing.

However, we do not claim that there are no distinctions between the friends, public, and work contact recipients. People are more comfortable sharing certain data types with certain human recipients. For instance, participants were significantly more uncomfortable sharing if they were lying, nervous, or stressed to work contacts compared to the rest of the data recipients. We believe a more fine-grained study is required to measure these effects accurately, although we have insight into the existence of these trends. Table \ref{all-vur} shows the complete VURs and rankings of all data types by recipient.

@@

\noindent
\textit{P1256: ``They can be hacked and then your security will be compromised.''}
\end{quotation}

Other common concerns included being unaware of what the device is collecting, doing, or which information it is using (Being Unaware), long-term health effects caused from wearing the device such as cancer from EM waves (Health), and safety hazards from wearing the device, such as distractions that cause car accidents (Safety).

@@

\noindent
\textit{P1038: ``Becoming distracted by the devices while doing other activities that require concentration such as driving.''}

@@

Interestingly, a few participants were concerned with resulting changes in social behaviors, such as dependence on devices or spending less time with loved ones (Social Impact).

\begin{quotation}
\noindent
\textit{P1425: I think the biggest risk is how they may effect society as a whole... a wearable technology that's always on and available may push things even further to the point where people spend less time actually interacting with loved ones, and applying their own critical thinking in certain situations, instead always relying on their devices.}
\end{quotation}

The landscape of users' perceived risks associated with wearables is broad, encompassing concepts outside of privacy and security. We therefore hope this motivates other researchers to investigate these other risks.

@@

While we initially observed an effect on VURs based on whether or not participants claimed to already own wearables (57.0\% vs. 60.8\%, respectively; Mann-Whitney $U=202,896$, $p\textless0.032$), this difference did not remain significant upon correcting for multiple testing (Bonferroni corrected $\alpha=0.01$). The effect of a participant's gender also did not remain significant upon correcting for multiple testing. We observed no correlation between a participant's education level and VUR.

@@

While the first three categories are self-explanatory, the latter three categories are based on user characteristics. We defined {\it behavioral information} as observations about the user's activities; {\it biometric information} as measurements of the user's body; and {\it demographic information} as non-biometric information about the user's traits. The risks for data types fell into five categories:

@@

Although we presented our participants with a prompt illustrating all the benefits of a wearable device, the questions we asked isolated the risk from the benefits of sharing data with a particular recipient. We sacrificed context due to the complexity of the question necessary for the participant to answer correctly. Users are willing to tolerate risks if there is enough benefit associated with that risk. We do believe that since all of our questions were out of context, our study does represent what data people unambiguously would like to be private and secure, and is accurate for measuring user perceptions of wearables.

@@

\section{Related Work}
\noindent {\bfseries User Perceptions} 
While risk communication for the physical world has been examined for several decades (e.g.,~\cite{slovic2000perception, Fischhoff,Morgan2001}), research into effectively communicating computer-based risks has only recently been researched. For example, Garg {\it et al.}\ and Blythe {\it et al.}\ show that computer-based risk communication should employ some degree of demographic targeting due to varying perceptions and abilities that correlate with demographic factors~\cite{Garg2012,Blythe2011}. This work is likely applicable to wearable computing risk communication, a more thorough understanding of users'  information risk acceptance in this domain is warranted prior to examining risk communication.

@@

One limitation of user perceptions is that people do not always have enough information to make privacy-sensitive decisions. Even if users did have this information, it has been shown that users often trade off long-term privacy for short-term benefits~\cite{acquisti2005privacy}. Furthermore, actual behavior may deviate from stated privacy preferences~\cite{spiekermann2001privacy}. However, understanding user concerns is a necessary first step not only for risk communication, but preventative measures against breaches of privacy and security in a new threat landscape.

\noindent {\bfseries Ubiquitous Sensing} We are rapidly moving towards a world of ubiquitous sensing and data capture, with ensuing privacy challenges~\cite{abowd2000charting,palen2003unpacking,camp2000internet}. Roesner {\it et al.}\ urge the community to address potential concerns for wearable devices before the technologies become widespread~\cite{roesner2014security} and explore the unique problems present in terms of law and policy~\cite{roesner2014augmented}. Privacy preservation research in this age of ubiquitous sensing include frameworks to design for privacy~\cite{bellotti1993design,camp2003designing,langheinrich2001privacy}, protocols for anonymous communication~\cite{cornelius2008anonysense}, evaluation metrics for privacy~\cite{scholtz2004toward}, and privacy models~\cite{hong2004privacy, jiang2002approximate}. Our work aims to guide these efforts with an insight into user acceptance of common information risks.

\noindent {\bfseries Lessons from Smartphones} Not long ago, smartphones revolutionized applications' access to  data. While this tends to benefit users, users often do not think of the privacy implications. Many unresolved concerns, such as the opaqueness that prevents users from fully understanding how applications are using their data or rogue applications inappropriately accessing data~\cite{1_kane_2010, zhou2011taming}, are applicable to wearables.

Felt {\it et al.} previously studied the security concerns of smartphone users through large-scale online survey~\cite{Felt}. Their survey asked 3,115 smartphone users about 99 risk scenarios. Participants were asked how upset they would be if a certain action occurred without their permission. Participants rated each situation on a Likert scale ranging from ``indifferent (1)'' to ``very upset (5).'' Our methodology closely follows that study, but with scenarios chosen to shed light on the security and privacy risks of wearable devices.

@@

\newpage
\input{tex-inputs/appendixes.tex}
\input{tex-inputs/VURcombined.tex}
